---
title: "Week 06 Quarto Homework"
author: "Timothy Li"
format: 
  html:
    code-fold: true
    code-summary: "Show Code"
    embed-resources: true
date: "`r Sys.Date()`"
editor: visual
execute:
  warning: false
  message: false
  eval: true
theme: zephyr
---

# I'm creating a **plot** and **table** with the Chemistry Data set from Silbiger et al. 2020

### The first step is loading the libraries and reading in the data

```{r}
### Load in Libraries #####

library(tidyverse)
library(here)
library(kableExtra)

##### Read in Data #####

ChemData <- read_csv(here("Week_06", "Data", "chemicaldata_maunalua.csv"))
glimpse(ChemData)

```

### The second step is cleaning up the data!

```{r}
##### Clean Data ##### 

ChemData_clean <- ChemData |>
  drop_na() |>            # Remove all NAs
  separate_wider_delim(cols  = Tide_time, # Separate the Tide_time column into 2 separate columns
                       delim = "_",
                       names = c("Tide", "Time"))

glimpse(ChemData_clean)
```

### The third step is determining what we want to plot and plotting it! I'm choosing to focus on the relationship between *Phosphate* and *Silicate* concentrations in reference to what *Season* it is and what *Time* of day it is.

```{r}
#| label: fig-phosphate-silicate
#| fig-cap: "The Relationship Between Silicate Concentrations (ppm) and Phosphate Concentrations (ppm) During Different Seasons at Day and Night"
#| fig-cap-location: bottom
#| out-width: "90%"
#| fig-width: 10
#| fig-height: 6
#| fig-align: "center"

phosphate_silicate_plot <- ggplot(data = ChemData_clean,
                           mapping = aes(x = Silicate,
                                         y = Phosphate,
                                         color = interaction(Season, Time) # Interaction lets me have separate colors for each grided plot
                                         )) +
  geom_point(size = 1, alpha = 0.5) +
  facet_grid(Time ~ Season, # Grid separated by time and season (4 total plots)
             labeller = labeller(
               Season = c("FALL" = "Fall", "SPRING" = "Spring"), # Labeller changes the label names 
               Time = c("Day" = "Daytime", "Night" = "Nighttime")
             ))  + 
  scale_fill_viridis_d(option = "D") + # Color-blind friendly scale
  geom_smooth(method = "lm", linewidth = .75, color = "grey30") +     
  labs(
       x = "Silicate Concentration (ppm)", y = "Phosphate Concentration (ppm)",
       caption = "Silbiger et al. 2020 Proceedings of the Royal Society: B"
  ) +
  theme_bw() +
  theme(
    plot.title = element_text(size = 13, hjust = .5, face = "bold"),
    axis.title.x = element_text(size = 10, face = "bold"),
    axis.title.y = element_text(size = 10, face = "bold"),
    plot.caption = element_text(size = 5),
    strip.text = element_text(size = 10, face = "bold"),
    axis.text.x = element_text(size = 7),
    axis.text.y = element_text(size = 7),
    legend.position = "none"
  )

phosphate_silicate_plot
```

### The final step is identifying some important takeaways from the plots. I also want to make a table of the important summary statistics to identify some of these takeaways much easier!

```{r}
#| label: tbl-phosphate-silicate
#| tbl-cap: "Summary Statistics for Phosphate (ppm) and Silicate (ppm) Concentrations"
#| tbl-cap-location: bottom

##### Pivot into long format and Calculate Summary Statistics #####

ChemData_summary <- ChemData_clean |>
  pivot_longer(cols = Phosphate:Silicate, # Making the dataset longer
               names_to  = "Variables",         
               values_to = "Values") |>
  group_by(Season, Time, Variables) |> # Summarizing for Variable and Zone
  summarise(Mean = mean(Values, na.rm = TRUE),
            Variability  = var(Values,  na.rm = TRUE),
            "Standard Deviation" = sd(Values, na.rm = TRUE)) |> # Finding mean, variability, and SD
  write_csv(here("Week_06", "Outputs", "HW06_summary.csv")) |>
  kbl() |> # Creates the table
  kable_classic() |> # Table theme
  kable_styling(full_width = FALSE) # Makes sure the table isn't too wide

ChemData_summary

```

### Key Takeaways from the plots and summary statistics:

- All 4 plots have relatively strong and positive correlations

- Points are heavily concentrated at lower levels of Phosphate and Silicate

- Daytime in the Spring and Nightime in the Fall have Silicate concentrations above 150 ppm

  - These could potentially be outliers which could be tested using IQR

- Daytime in Spring has the highest mean Silicate and Phosphate concentrations (ppm)

- `r max(ChemData_clean$Silicate)` is the max Silicate concentration (ppm)

- `r max(ChemData_clean$Phosphate)` is the max Phosphate concentration (ppm)

### That is all. Thanks!
